---
title: "Synthetic Illustration with the code"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
  pdf_document:
    toc: true
    latex_engine: xelatex
---

This document walks through the article's Empirical Illustration with
the code for each step. It runs on the table `import-data.R` names,
which as shipped is the synthetic low-consolidation file, standing in
for the case study. Every value below comes from that file. Knit it in
RStudio, or render it from the repository root with
`rmarkdown::render("synthetic_walkthrough.Rmd", output_dir = "output")`.
To see the same walkthrough on the moderate or high file, or on your
own table, change `data_path` and knit again. The scripts in `methods/`
run the same steps one section at a time. This document fits the local
types itself and does not write methods/03's run of record, which
methods/04 and 05 read; run methods/03 for that.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(arules)
library(poLCA)
library(pscl)
library(vcd)
for (f in list.files("R_functions", pattern = "[.]R$", full.names = TRUE)) source(f)
source("import-data.R")
knitr::opts_chunk$set(message = TRUE, warning = FALSE)
data <- oc_load_data()
```

The check the entry file runs on its own, here on the loaded table:

```{r}
check_data(data, oc_config)
```

# An initial assessment of organizational consolidation

The article's initial assessment uses two complementary strategies:
pairwise associations between each ascriptive status and each local
criterion, and mutual information, which captures nonlinear
dependencies. Each pair is measured with the statistic matched to its
levels of measurement and reported as a 0–1 magnitude (Figure 4).

```{r}
set.seed(1234)
cons <- measure_consolidation(data,
                              ascriptive = oc_config$ascriptive,
                              criteria   = oc_config$criteria)
print(cons)
```

```{r fig-initial-assessment, fig.width=8.2, fig.height=4.5}
print(plot_initial_assessment(cons, data))
```

# Surfacing salient attributes: association rule mining

Association rule mining identifies if–then rules, if [Trait A + Trait
B] then [Status Position X], that show how attributes convey meaning in
relation. Its payoff is twofold: which attributes function as standalone
cues and which gain meaning only in bundles, and where the same
attribute carries different meanings in different pairings. Each
nomination is one transaction, and the article's filter chain keeps
rules with support of at least 0.05, confidence of at least 0.50, and
lift above 1 (Appendix B).

```{r}
transactions <- noms_to_transactions(
  data,
  attributes = c(oc_config$ascriptive, oc_config$criteria),
  outcomes   = oc_config$outcomes,
  unit       = oc_config$arm_unit)
rules <- apriori(
  transactions,
  parameter  = list(supp = 0.05, conf = 0.50, target = "rules", minlen = 2),
  appearance = list(lhs = setdiff(colnames(transactions), oc_config$outcomes),
                    rhs = oc_config$outcomes),
  control    = list(verbose = FALSE))
rules <- add_rule_measures(rules, transactions)
rules <- subset(rules, lift > 1)
cat(length(rules), "rules survive the filter chain.\n\n")
for (o in oc_config$outcomes) {
  cat("--", toupper(o), "--\n")
  arules::inspect(head(rules_for(rules, o), 5), linebreak = FALSE)
  cat("\n")
}
```

The rule network (Figure 6): attribute values feed rule nodes, sized by
conviction, and each rule points at the status position it predicts.
`max_rules_per_position` in `import-data.R` caps what is drawn, as in
methods/02.

```{r fig-rule-network, fig.width=9, fig.height=6.5}
plot_rules <- rules_to_draw(rules, oc_config$outcomes,
                            cap = oc_config$max_rules_per_position)
print(oc_provenance(plot_rule_network(
  plot_rules, outcomes = oc_config$outcomes,
  labels = oc_dummy_labels(data, c(oc_config$ascriptive, oc_config$criteria),
                           oc_config$outcomes, oc_config$labels))))
```

# Identifying local types: latent class analysis

Latent class analysis operationalizes local types, the prevalent
clusters of local criteria that co-occur within an organization's
population. Only the local criteria enter the analysis; race and gender
stay outside it, so that whether the types overlap with or cut across
the ascriptive statuses can be tested afterward. The class count is
chosen from the fit statistics of Appendix C (Table C1); `import-data.R`
sets `n_types = 4` for the synthetic file.

```{r}
set.seed(1234)
fit_stats <- lca_fit_stats(data, oc_config$criteria, k_range = 1:6)
print(as_table_c1(fit_stats), digits = 4, row.names = FALSE)
```

```{r fig-class-sweep, fig.width=6.5, fig.height=4}
print(oc_provenance(plot_class_count_sweep(fit_stats)))
```

```{r}
criteria <- oc_config$criteria
manifest <- as.data.frame(lapply(data[criteria], as_lca_codes))   # each criterion coded 1..K
f <- as.formula(paste("cbind(", paste(criteria, collapse = ", "), ") ~ 1"))
set.seed(1234)
LCA <- poLCA(f, manifest, nclass = oc_config$n_types, nrep = 10,
             maxiter = 5000, tol = 1e-10, na.rm = FALSE, calc.se = TRUE,
             graphs = FALSE, verbose = FALSE)
LCA$assignment <- factor(LCA$predclass, levels = seq_len(oc_config$n_types))
data$local_type <- LCA$assignment
cat("Class shares:", paste(round(prop.table(table(LCA$assignment)), 2), collapse = "  "), "\n")
cat("Entropy R^2:", round(entropy.R2(LCA), 2), "\n")
```

Whether local types overlap with or cut across ascriptive statuses is
the most direct assessment of organizational consolidation: Cramér's V
between type and status, and the mosaic of Figure 8, where color marks
the cells that depart from independence.

```{r}
for (a in oc_config$ascriptive) {
  cat(sprintf("Cramér's V, local_type x %s: %.3f\n",
              a, cramers_v(data$local_type, data[[a]])))
}
```

```{r fig-mosaic, fig.width=6.5, fig.height=4.5}
draw_type_mosaic(data, oc_config$ascriptive[1])
```

# Local types as regression inputs

Local types also serve as inputs for regression: using the LCA-derived
profiles as predictors preserves the relational structure rule mining
surfaced, rather than decomposing it into independent effects. The
models are zero-inflated Poisson regressions of nomination counts on
race, gender, their intersections, the local types, and the controls
(Appendix D). Local types enter by proportional assignment, each
member's posterior class probabilities. Appendix D's tables compare
three models per status position: ascriptive statuses alone, local
types alone, and both combined.

```{r}
type_probs <- setNames(as.data.frame(LCA$posterior[, -1, drop = FALSE]),
                       paste0("type_prob", seq_len(oc_config$n_types)[-1]))
data <- cbind(data, type_probs)
ascriptive    <- oc_config$ascriptive
intersections <- if (length(ascriptive) >= 2) combn(ascriptive, 2, paste, collapse = ":") else character(0)
controls      <- oc_config$controls
focal         <- oc_config$focal_outcome
dist          <- oc_config$count_model
options(contrasts = c("contr.treatment", "contr.treatment"))   # as Appendix D coded them
data <- dummy_coded(data, c(ascriptive, intersections, names(type_probs), controls))
three <- list(
  "Ascrip."    = zeroinfl(reformulate(c(ascriptive, intersections, controls), focal),
                          data = data, dist = dist, link = "logit"),
  "Local Type" = zeroinfl(reformulate(c(names(type_probs), controls), focal),
                          data = data, dist = dist, link = "logit"),
  "Combined"   = zeroinfl(reformulate(c(ascriptive, intersections, names(type_probs), controls), focal),
                          data = data, dist = dist, link = "logit"))
knitr::kable(recognition_model_table(three, term_order = c(names(type_probs), ascriptive, controls)),
             row.names = FALSE)
```

The combined model's predicted nominations for each intersectional
profile are what the status hierarchy orders.

```{r}
profiles <- predicted_profiles(three$Combined, data,
                               profile_vars = c(ascriptive, "local_type")) %>%
  filter(n >= 3)
print(head(profiles[order(-profiles$predicted), ], 5), digits = 3)
```

# Mapping intersectionality: partial order set modeling

Posets show how context-specific combinations of attributes are ordered
relative to the organization's ideal actor, the intersectional profile
that receives maximum recognition. The article focuses on "respected".
Profiles are arranged by how many high-value characteristics they share
with the ideal, and the predicted nomination values set each profile's
vertical position (Figure 9). The construction runs on the combined
model above, one step per call, as `methods/05` does.

Every position, each ascriptive status by local type, is valued by
prediction: the model predicts each member's recognition as if they
held that position, with their own controls, and the average is the
position's value.

```{r}
data$local_type <- label_local_types(LCA$assignment, oc_config$type_labels)
dimensions <- c(ascriptive, "local_type")
positions  <- predict_positions(three$Combined, data, dimensions, names(type_probs))
print(head(positions, 5), digits = 3)
```

The ideal actor is chosen among positions with at least three members,
by the highest lower confidence bound; `rule = "max"` takes the highest
point value, the article's rule.

```{r}
ideal <- choose_ideal(positions, rule = "conf")
cat("Ideal actor:", paste(unlist(ideal$full), collapse = " / "), "\n")
```

Each dimension is then reduced to the ideal's level versus the rest,
and each cell of the cube takes the pooled prediction of its positions.

```{r}
cube <- collapse_to_cube(three$Combined, positions, data, ideal)
print(cube, digits = 3)
```

Two drawings follow: the formal structure, then the same cube with
each cell at its predicted recognition.

```{r fig-hierarchy-formal, fig.width=11, fig.height=6}
formal <- status_hierarchy_poset(cube, layout = "formal", label_style = "match")
print(oc_provenance(formal$plot))
```

```{r fig-hierarchy-values, fig.width=9, fig.height=6}
values <- status_hierarchy_poset(cube, layout = "outcome", label_style = "match",
                                 axis_limits = oc_config$poset_axis_limits)
print(oc_provenance(values$plot +
        labs(y = sprintf("Expected %s nominations", focal))))
print(values$shape_metrics, digits = 2)
```

The article's diagnostic step after Figure 9: among members who share
the ideal's local type, how many in each ascriptive profile were
recognized, and how many would have been had that type been rewarded at
the ideal's rate.

```{r}
print(unrecognized_members(three$Combined, data, ideal), digits = 3, row.names = FALSE)
```
